On type cones of $\b{g}$-vector fans

### 2019-11-11 thirdPaper/FPSAC/typeConeAssociahedraFPSAC.tex
keywords
resume
corrections + pictures FPSAC

%% define your authors in the usual way
%% use \addressmark{1}, \addressmark{2} etc for the institutions, and use \thanks{} for contact details
\author[Arnau Padrol, Yann Palu, Vincent Pilaud and Pierre-Guy Plamondon]{
	Arnau Padrol\thanks{\href{mailto:arnau.padrol@imj-prg.fr}{arnau.padrol@imj-prg.fr}. Supp. by ANR grant CAPPS~(17\,CE40\,0018).}\addressmark{1},
	\stepcounter{footnote}
	Yann Palu\thanks{\href{mailto:yann.palu@u-picardie.fr}{yann.palu@u-picardie.fr}. Supp. by ANR grant SC3A~(15\,CE40\,0004\,01).}\addressmark{2}, \\
	Vincent Pilaud\thanks{\href{mailto:vincent.pilaud@lix.polytechnique.fr}{vincent.pilaud@lix.polytechnique.fr}. Supp. by ANR SC3A~15\,CE40\,0004\,01 and CAPPS~17\,CE40\,0018.}\addressmark{3}
	\and
	Pierre-Guy Plamondon\thanks{\href{mailto:pierre-guy.plamondon@math.u-psud.fr}{pierre-guy.plamondon@math.u-psud.fr}. Supp. by ANR grant SC3A~(15\,CE40\,0004\,01).}\addressmark{4},
}

@@

%% put your English abstract here, or comment this out if you don't have one yet
%% please don't use custom commands in your abstract / resume, as these will be displayed online
%% likewise for citations -- please don't use \cite, and instead write out your citation as something like (author year)
\abstract{
We study the type cone (\ie the space of all polytopal realizations) of $\b{g}$-vector fans of finite type cluster-like complexes (finite type cluster complexes, non-kissing complexes of gentle algebra, and graphical nested complexes). We show that this cone is often simplicial, which explains an elegant ``kinematic'' construction of the associahedron as a section of a high dimensional positive orthant by certain affine subspaces parametrized by a small dimensional positive orthant.
}

%% put your French abstract here, or comment this out if you don't have one
\resume{
Nous \'etudions le type cone (\ie l'espace de toutes les r\'ealisations polytopales) des \'eventails de $\b{g}$-vecteurs de complexes de type amassés (complexes amassés de type fini, complexes platoniques d'alg\`ebres aimables, et complexes emboit\'es graphi\-ques). Nous montrons que ce cone est souvent simplicial, ce qui explique une \'el\'egante construction ``cin\'ematique'' de l'associa\`edre comme section d'un orthant positif de grande dimension par certains sous-espaces affines param\`etrés par un orthant positif de petite dimension.
}

%% put your keywords here, or comment this out if you don't have them yet
\keywords{Fans $\cdot$ cluster algebras $\cdot$ non-kissing complex $\cdot$ graph associahedra}

@@

Besides revisiting the construction of~\cite{ArkaniHamedBaiHeYan, BazierMatteDouvilleMousavandThomasYildirim} and extending it to any initial seed, our type cone approach is also successful when applied to the $\b{g}$-vector fans of other families of generalizations of the associahedron.
In the present paper, we explore specifically the non-kissing complexes of gentle algebras introduced in~\cite{PaluPilaudPlamondon-nonkissing} (see \cref{subsec:gentleAssociahedra}) and the graphical nested complexes studied in~\cite{CarrDevadoss, Postnikov, FeichtnerSturmfels, Zelevinsky} (see \cref{subsec:graphAssociahedra}).

@@

Fix an essential complete simplicial fan\footnote{A \defn{polyhedral cone} is the positive span of finitely many vectors or equivalently, the intersection of finitely many closed linear half-spaces. The \defn{faces} of a cone are its intersections with its supporting hyperplanes. A \defn{fan}~$\Fan$ is a set of polyhedral cones such that any face of a cone of~$\Fan$ belongs to~$\Fan$, and any two cones of~$\Fan$ intersect along a face of both. A fan is \defn{essential} if the intersection of its cones is the origin, \defn{complete} if the union of its cones covers~$\R^n$, and \defn{simplicial} if all its cones are generated by dimension many rays.}~$\Fan$ in~$\R^n$. Let~$\b{G}$ the $N \times n$-matrix whose rows are (representative vectors of) the rays of~$\Fan$ and let~$\b{K}$ be a $(N-n) \times N$-matrix that spans the left kernel of~$\b{G}$ (\ie $\b{K}\b{G} = 0$). For any height vector~$\b{h} \in \R^N$, we define the polytope\footnote{A \defn{polytope} is the convex hull of finitely many points or equivalently, a bounded intersection of finitely many closed affine half-spaces.}
\(
P_\b{h} \eqdef \set{\b{x} \in \R^n}{\b{G}\b{x} \le \b{h}}.
\)
Unfortunately, the normal fan\footnote{The \defn{normal cone} of a face~$F$ of a polytope~$P$ is the cone generated by the normal vectors of the facets of~$P$ containing~$F$. The \defn{normal fan} of~$P$ is the set of normal cones of all its faces.} of this polytope~$P_\b{h}$ does not always coincide with the fan~$\Fan$ we started from. The following statement gives a simple characterization of the height vectors~$h$ for which it is the case. It is a reformulation of regularity of triangulations of vector configurations, introduced in the theory of secondary polytopes~\cite{GelfandKapranovZelevinsky}.
We present here a convenient formulation from~\cite{ChapotonFominZelevinsky}.

@@

\begin{example}
\label{exm:typeCone}
\enlargethispage{.1cm}
Consider the $2$-dimensional fan~$\Fan$ depicted in \cref{fig:typeCone} (left).
% (we will encounter this fan again in \cref{exm:typeConeCoarsenings,exm:typeConeMinkowskiSum,exm:typeConeAsso}).
It has five rays labeled~$1, \dots, 5$ and five maximal cones labeled~$a,\dots, e$.
Its type cone~$\typeCone(\Fan)$ lies in~$\R^5$, but has a $2$-dimensional linearity space.
The five pairs of adjacent maximal cones of~$\Fan$ give rise to following five defining inequalities for~$\typeCone(\Fan)$:
\begin{gather*}
H^>_{ab}: \; \b{h}_1 + \b{h}_3 - \b{h}_2 > 0
\qquad
H^>_{bc}: \; \b{h}_2 + \b{h}_4 - \b{h}_3 > 0
\qquad
H^>_{cd}: \; \b{h}_3 + \b{h}_5 - \b{h}_4 > 0
\\
H^>_{de}: \; \b{h}_1 + \b{h}_4 > 0
\qquad
H^>_{ae}: \; \b{h}_2 + \b{h}_5 > 0,
\end{gather*}
where $H^>_{ij}$ denotes the halfspace defined by the inequality corresponding to the two adjacent maximal cones~$i$ and~$j$.
%Note that the inequalities~$H^>_{ae}$ and~$H^>_{de}$ are redundant, while the other three define facets, so that the extremal adjacent pairs are~$(a,b)$, $(b,c)$ and~$(c,d)$.
For example, the height vector~$(\nicefrac{1}{2}, \nicefrac{3}{4}, \nicefrac{5}{4}, 1, \nicefrac{5}{4})$ belongs to the type cone~$\typeCone(\Fan)$, and the corresponding polytope is represented in \cref{fig:typeCone} (middle).
To represent this type cone, we slice it with a transversal $2$-dimensional affine space. This is the red triangle depicted in \cref{fig:typeCone} (right).

@@

\begin{remark}
All fans considered in \cref{sec:gvectorFans} have the \defn{unique exchange relation property}: the linear dependence ${\displaystyle\sum}_{\ray[s] \in \rays \cup \rays'} \coefficient[{\ray[s]}][\rays][\rays'] \, \ray[s] = 0$ only depends on the exchanged rays~$\ray[r]$ and~$\ray[r]'$, not on the maximal adjacent cones~$\rays$ and~$\rays'$. We then write~$\coefficient[{\ray[s]}][\ray][\ray']$ instead of~$\coefficient[{\ray[s]}][\rays][\rays']$ and we call \defn{extremal adjacent pair} of~$\Fan$ the pairs of exchangeable rays~$\{\ray, \ray'\}$ of~$\Fan$ such that the corresponding inequality ${\displaystyle\sum}_{\ray[s] \in \rays \cup \rays'} \coefficient[{\ray[s]}][\ray][\ray'] \, \b{h}_{\ray[s]} > 0$ defines a facet of~$\typeCone(\Fan)$.

@@

\begin{proposition}
\label{prop:alternativePolytopalRealization}
The affine map~$\Psi: \R^n \to \R^N$ defined by~$\Psi(\b{x}) = \b{h} - G\b{x}$ sends the polytope~$P_\b{h} \eqdef \set{\b{x} \in \R^n}{\b{G}\b{x} \le \b{h}}$ to the polytope~$Q_\b{h} \eqdef \set{\b{z} \in \R^N}{\b{K}\b{z} = \b{K}\b{h} \text{ and } \b{z} \ge 0}$.
\end{proposition}

@@

We now study the type cones of the normal fans of three families generalizing the associahedron constructed in~\cite{Loday}: the generalized associahedra of finite type cluster algebras~\cite{FominZelevinsky-ClusterAlgebrasI, FominZelevinsky-ClusterAlgebrasII, FominZelevinsky-ClusterAlgebrasIV, HohlwegLangeThomas, HohlwegPilaudStella}, the gentle associahedra~\cite{PaluPilaudPlamondon-nonkissing}, and the graph associahedra~\mbox{\cite{CarrDevadoss, Postnikov, FeichtnerSturmfels, Zelevinsky}}.
%All these families contain the classical associahedra~$\Asso[n]$ constructed in~\cite{Loday}.

@@

%We start with the classical associahedron.
A $n$-dimensional \defn{associahedron} is a polytope whose face lattice is the reverse inclusion lattice of dissections (\ie sets of pairwise non-crossing diagonals) of a convex~$(n+3)$-gon.
In particular, its vertices correspond to triangulations of the $(n+3)$-gon and its facets correspond to internal diagonals of the $(n+3)$-gon.
Let~$X(n) \eqdef \set{(a,b)}{0 \le a < b \le n+2}$ (resp.~$Y(n) \eqdef \set{(a,b)}{0 \le a < b-2 \le n}$) denote all (resp.~the internal) diagonals.
The \defn{$\b{g}$-vector} of a diagonal $(a,b) \in X(n)$ is ${\gvector{a,b} \eqdef (n+1) {\displaystyle\sum}_{\,a < \ell < b} \b{e}_\ell - (b-a-1) {\displaystyle\sum}_{\,1 \le \ell \le n+1} \b{e}_\ell}$.
%Note that~$\gvector{a,b} = 0$ if~$(a,b)$ is not internal.
We set~$\gvector{D} \eqdef \set{\gvector{a,b}}{(a,b) \in D}$ for a dissection~$D$.
Recall that:
\begin{compactitem}
\item the set of cones
\(
\gvectorFan[n] \eqdef \bigset{\R_{\ge0} \, \gvectors{D}}{D \text{ dissection of the $(n+3)$-gon}}
\)
forms a complete simplicial fan.
See \cref{fig:lodayFans}\,(left and middle).

@@

\begin{figure}[t]
	\centerline{
	\raisebox{.5cm}{\begin{overpic}[scale=.5]{lodayFan2} \put(0,105){$\boxed{n=2}$} \end{overpic}}
	\hspace{.5cm}
	\begin{overpic}[scale=.5]{lodayFan3} \put(-10,93){$\boxed{n=3}$} \end{overpic}
	\hspace{.5cm}
	\raisebox{.35cm}{\begin{overpic}[scale=.5]{clusterFanA}
		\put(35,33){${\scriptstyle x_1}$}
		\put(47,54){${\scriptstyle x_2}$}
		\put(59,33){${\scriptstyle x_3}$}
		\put(70,37){$\frac{x_2 + x_3}{x_1}$}
		\put(22,18){$\frac{x_1 + x_3}{x_2}$}
		\put(29,62){$\frac{x_1 + x_2}{x_3}$}
		\put(80,57){$\frac{x_1 + x_2 + x_3}{x_1 x_3}$}
		\put(35,-7){$\frac{x_1 + x_2 + x_3}{x_1 x_2}$}
		\put(-10,57){$\frac{x_1 + x_2 + x_3}{x_2 x_3}$}
%		\put(34,49){$x_1$}
%		\put(57,61){$x_2$}
%		\put(57,36){$x_3$}
%		\put(68,33){$\frac{x_2 + x_3}{x_1}$}
%		\put(17,36){$\frac{x_1 + x_3}{x_2}$}
%		\put(44,74){$\frac{x_1 + x_2}{x_3}$}
%		\put(90,49){$\frac{x_1 + x_2 + x_3}{x_1 x_3}$}
%		\put(7,12){$\frac{x_1 + x_2 + x_3}{x_1 x_2}$}
%		\put(7,84){$\frac{x_1 + x_2 + x_3}{x_2 x_3}$}
	\end{overpic}}
	}
	\caption{The fans~$\gvectorFan[2]$, $\gvectorFan[3]$ and~$\gvectorFan[\B_\circ]$. The two right $3$-dimensional fans are intersected with the sphere and stereographically projected from the direction~$(-1,-1,-1)$.}
	\label{fig:lodayFans}
	\vspace*{-.2cm}
\end{figure}

@@

For~$\{x,x'\} \in \meshes$ and~$y \in \variables$, we denote by~$\coefficient[y][x][x']$ the coefficient of~$\gvectorFull{\B_\circ}{y}$ in the linear dependence among the $\b{g}$-vectors~$\gvectorFull{\B_\circ}{x}$ and~$\gvectorFull{\B_\circ}{x'}$. In other words, by \cref{lem:dependenceMesh}, if~${(\B, \cluster)}$ and~${(\B', \cluster')}$ are for two adjacent seeds with~${\cluster \ssm \{x\} = \cluster' \ssm \{x'\}}$, we have~$\coefficient[y][x][x'] = |b_{xy}|$ for~$y \in \cluster \cap \cluster'$ and~$\coefficient[y][x][x'] = 0$ otherwise.
The next statement is proved in~\cite[Sect.~3]{PadrolPaluPilaudPlamondon} using techniques from the quiver representation theory.

@@

%We derive from \cref{thm:meshMutations} the following description of the facets of the type cone of the cluster fan.
Using \cref{thm:meshMutations} as a blackbox, we describe the facets of the type cone. See \cref{fig:labelFacetDefiningInequalititiesAsso}. %, and thus all polytopal realizations, of the cluster fans of finite type cluster algebras.

\begin{theorem}
\label{thm:extremalExchangeablePairsCA}
For any finite type exchange matrix~$\B_\circ$, the type cone~$\typeCone \big( \gvectorFan[\B_\circ] \big)$ is simplicial and the non-initial mesh mutations are the extremal adjacent pairs of the cluster fan~$\gvectorFan[\B_\circ]$.
\end{theorem}

@@

A \defn{substring} of a walk~$\smash{\omega = \alpha_1^{\varepsilon_1} \cdots \alpha_\ell^{\varepsilon_\ell}}$ of~$\quiver$ is a string~$\smash{\sigma = \alpha_{i+1}^{\varepsilon_{i+1}} \cdots \alpha_{j-1}^{\varepsilon_{j-1}}}$ of~$\quiver$ for some indices~$1 \le i < j \le \ell$.
We say that the substring~$\smash{\sigma = \alpha_{i+1}^{\varepsilon_{i+1}} \cdots \alpha_{j-1}^{\varepsilon_{j-1}}}$ is \defn{at the bottom} (resp.~\defn{on top}) of the walk~$\omega = \alpha_1^{\varepsilon_1} \cdots \alpha_\ell^{\varepsilon_\ell}$ if~$\varepsilon_i = 1$ and~$\varepsilon_j = -1$ (resp.~if~$\varepsilon_i = -1$ and~$\varepsilon_j = 1$).
In other words the two arrows of~$\omega$ incident to the endpoints of~$\sigma$ point towards~$\sigma$ (resp.~outwards from~$\sigma$).
We denote by~$\Sigma_\bottom(\omega)$ and~$\Sigma_\top(\omega)$ the sets of bottom and top substrings of~$\omega$ respectively.

@@

	\caption{The facet-defining inequalities of the type cone~$\typeCone \big( \gvectorFan[\quiver] \big)$ for the non-kissing fan of \cref{fig:nonkissingFans}\,(right). The conventions are similar as in \cref{fig:labelFacetDefiningInequalititiesAsso}.}

@@

\begin{definition}
\label{def:brick2acyclic}
A gentle quiver~$\quiver$ is called:
\begin{compactitem}
\item \defn{brick} if any (non necessarily oriented) cycle of~$\quiver$ contains at least two relations in~$I$. %, or equivalently, no walk on~$\quiver$ is self-kissing.
\item \defn{$2$-acyclic} if it contains no cycle of length~$2$.
\end{compactitem}
\end{definition}

@@

We now consider the graph associahedra studied in~\cite{CarrDevadoss, Postnikov, FeichtnerSturmfels, Zelevinsky}.
Again, we briefly recall all definitions needed here as they are purely combinatorial.

>>>>

%% define your authors in the usual way
%% use \addressmark{1}, \addressmark{2} etc for the institutions, and use \thanks{} for contact details
\author[Arnau Padrol, Yann Palu, Vincent Pilaud and Pierre-Guy Plamondon]{
	Arnau Padrol\thanks{\href{mailto:arnau.padrol@imj-prg.fr}{arnau.padrol@imj-prg.fr}. Supp. by ANR grant CAPPS~(17\,CE40\,0018).}\addressmark{1},
	\stepcounter{footnote}%
	Yann Palu\thanks{\href{mailto:yann.palu@u-picardie.fr}{yann.palu@u-picardie.fr}. Supp. by ANR grant SC3A~(15\,CE40\,0004\,01).}\addressmark{2}, \\
	Vincent Pilaud\thanks{\href{mailto:vincent.pilaud@lix.polytechnique.fr}{vincent.pilaud@lix.polytechnique.fr}. Supp. by ANR SC3A~15\,CE40\,0004\,01 and CAPPS~17\,CE40\,0018.}\addressmark{3}
	\and
	Pierre-Guy Plamondon\thanks{\href{mailto:pierre-guy.plamondon@math.u-psud.fr}{pierre-guy.plamondon@math.u-psud.fr}. Supp. by ANR grant SC3A~(15\,CE40\,0004\,01).}\addressmark{4},
}

@@

%% put your English abstract here, or comment this out if you don't have one yet
%% please don't use custom commands in your abstract / resume, as these will be displayed online
%% likewise for citations -- please don't use \cite, and instead write out your citation as something like (author year)
\abstract{
We study the type cone (\ie the space of all polytopal realizations) of $\b{g}$-vector fans of finite type cluster-like complexes (finite type cluster complexes, non-kissing complexes of gentle algebra, and graphical nested complexes). We show that this cone is often simplicial, which explains an elegant ``kinematic'' construction of the associahedron as a section of a high dimensional positive orthant by certain affine subspaces parametrized by a low dimensional positive orthant.
}

%% put your French abstract here, or comment this out if you don't have one
\resumetitle{R\'esum\'e}
\resume{
Nous \'etudions le type cone (\ie l'espace de toutes les r\'ealisations polytopales) des \'eventails de $\b{g}$-vecteurs de complexes de type amassés (complexes amassés de type fini, complexes platoniques d'alg\`ebres aimables, et complexes emboit\'es graphi\-ques). Nous montrons que ce cone est souvent simplicial, ce qui explique une \'el\'egante construction ``cin\'ematique'' de l'associa\`edre comme section d'un orthant positif de haute dimension par certains sous-espaces affines param\`etrés par un orthant positif de basse dimension.
}

%% put your keywords here, or comment this out if you don't have them yet
\keywords{Fans $\cdot$ cluster algebras $\cdot$ non-kissing complexes $\cdot$ graph associahedra}

@@

Besides revisiting the construction of~\cite{ArkaniHamedBaiHeYan, BazierMatteDouvilleMousavandThomasYildirim} and extending it to any initial seed, our type cone approach is also successful when applied to the $\b{g}$-vector fans of other families of generalizations of the associahedron.
In the present paper, we explore specifically the non-kissing complexes of gentle algebras introduced in~\cite{PaluPilaudPlamondon-nonkissing} (see \cref{subsec:gentleAssociahedra}) and the graphical nested complexes studied in~\cite{CarrDevadoss, FeichtnerSturmfels, Postnikov, Zelevinsky} (see \cref{subsec:graphAssociahedra}).

@@

Fix an essential complete simplicial fan\footnote{A \defn{polyhedral cone} is the positive span of finitely many vectors or equivalently, the intersection of finitely many closed linear half-spaces. The \defn{faces} of a cone are its intersections with its supporting hyperplanes. A \defn{fan}~$\Fan$ is a set of polyhedral cones such that any face of a cone of~$\Fan$ belongs to~$\Fan$, and any two cones of~$\Fan$ intersect along a face of both. A fan is \defn{essential} if the intersection of its cones is the origin, \defn{complete} if the union of its cones covers~$\R^n$, and \defn{simplicial} if all its cones are generated by dimension many rays.}~$\Fan$ in~$\R^n$. Let~$\b{G}$ be the $N \times n$-matrix whose rows are (representative vectors of) the rays of~$\Fan$ and let~$\b{K}$ be a $(N-n) \times N$-matrix that spans the left kernel of~$\b{G}$ (\ie $\b{K}\b{G} = 0$). For any height vector~$\b{h} \in \R^N$, we define the polytope\footnote{A \defn{polytope} is the convex hull of finitely many points or equivalently, a bounded intersection of finitely many closed affine half-spaces.}
\(
P_\b{h} \eqdef \set{\b{x} \in \R^n}{\b{G}\b{x} \le \b{h}}.
\)
Unfortunately, the normal fan\footnote{The \defn{normal cone} of a face~$F$ of a polytope~$P$ is the cone generated by the normal vectors of the facets of~$P$ containing~$F$. The \defn{normal fan} of~$P$ is the set of normal cones of all its faces.} of this polytope~$P_\b{h}$ does not always coincide with the fan~$\Fan$ we started from. The following statement gives a simple characterization of the height vectors~$h$ for which it is the case. It is a reformulation of regularity of triangulations of vector configurations, introduced in the theory of secondary polytopes~\cite{GelfandKapranovZelevinsky}.
We present here a convenient formulation from~\cite{ChapotonFominZelevinsky}.

@@

\begin{example}
\label{exm:typeCone}
\enlargethispage{.1cm}
Consider the $2$-dimensional fan~$\Fan$ depicted in \cref{fig:typeCone} (left).
% (we will encounter this fan again in \cref{exm:typeConeCoarsenings,exm:typeConeMinkowskiSum,exm:typeConeAsso}).
It has five rays labeled~$1, \dots, 5$ and five maximal cones labeled~$a,\dots, e$.
Its type cone~$\typeCone(\Fan)$ lies in~$\R^5$, but has a $2$-dimensional linearity space.
The five pairs of adjacent maximal cones of~$\Fan$ give rise to following five defining inequalities for~$\typeCone(\Fan)$:
\begin{gather*}
H^>_{ab}: \; \b{h}_1 + \b{h}_3 - \b{h}_2 > 0
\qquad
H^>_{bc}: \; \b{h}_2 + \b{h}_4 - \b{h}_3 > 0
\qquad
H^>_{cd}: \; \b{h}_3 + \b{h}_5 - \b{h}_4 > 0
\\
H^>_{de}: \; \b{h}_1 + \b{h}_4 > 0
\qquad
H^>_{ae}: \; \b{h}_2 + \b{h}_5 > 0,
\end{gather*}
where $H^>_{xy}$ denotes the halfspace defined by the inequality corresponding to the two adjacent maximal cones~$x$ and~$y$.
%Note that the inequalities~$H^>_{ae}$ and~$H^>_{de}$ are redundant, while the other three define facets, so that the extremal adjacent pairs are~$(a,b)$, $(b,c)$ and~$(c,d)$.
For example, the height vector~$(\nicefrac{1}{2}, \nicefrac{3}{4}, \nicefrac{5}{4}, 1, \nicefrac{5}{4})$ belongs to the type cone~$\typeCone(\Fan)$, and the corresponding polytope is represented in \cref{fig:typeCone} (middle).
To represent this type cone, we slice it with a transversal $2$-dimensional affine space. This is the red triangle depicted in \cref{fig:typeCone} (right).

@@

\begin{remark}
All fans considered in \cref{sec:gvectorFans} have the \defn{unique exchange relation property}: the linear dependence ${\displaystyle\sum}_{\ray[s] \in \rays \cup \rays'} \coefficient[{\ray[s]}][\rays][\rays'] \, \ray[s] = 0$ only depends on the exchanged rays~$\ray[r]$ and~$\ray[r]'$, not on the maximal adjacent cones~$\rays$ and~$\rays'$. We then write~$\coefficient[{\ray[s]}][\ray][\ray']$ instead of~$\coefficient[{\ray[s]}][\rays][\rays']$ and we call \defn{extremal exchangeable pairs} of~$\Fan$ the pairs of exchangeable rays~$\{\ray, \ray'\}$ of~$\Fan$ such that the corresponding inequality ${\displaystyle\sum}_{\ray[s] \in \rays \cup \rays'} \coefficient[{\ray[s]}][\ray][\ray'] \, \b{h}_{\ray[s]} > 0$ defines a facet of~$\typeCone(\Fan)$.

@@

\begin{proposition}
\label{prop:alternativePolytopalRealization}
The map~$\b{x} \in \R^n \mapsto \b{h} - G\b{x} \in \R^N$ is an affine transformation of the polytope~$P_\b{h} \eqdef \set{\b{x} \in \R^n}{\b{G}\b{x} \le \b{h}}$ to the polytope~$Q_\b{h} \eqdef \set{\b{z} \in \R^N}{\b{K}\b{z} = \b{K}\b{h} \text{ and } \b{z} \ge 0}$.
\end{proposition}

@@

We now study the type cones of the normal fans of three families generalizing the associahedron constructed in~\cite{Loday}: the generalized associahedra of finite type cluster algebras~\cite{FominZelevinsky-ClusterAlgebrasI, FominZelevinsky-ClusterAlgebrasII, FominZelevinsky-ClusterAlgebrasIV, HohlwegLangeThomas, HohlwegPilaudStella}, the gentle associahedra~\cite{PaluPilaudPlamondon-nonkissing}, and the graph associahedra~\mbox{\cite{CarrDevadoss, FeichtnerSturmfels, Postnikov, Zelevinsky}}.
%All these families contain the classical associahedra~$\Asso[n]$ constructed in~\cite{Loday}.

@@

%We start with the classical associahedron.
An $n$-dimensional \defn{associahedron} is a polytope whose face lattice is the reverse inclusion lattice of dissections (\ie sets of pairwise non-crossing diagonals) of a convex~$(n+3)$-gon.
In particular, its vertices correspond to triangulations of the $(n+3)$-gon and its facets correspond to internal diagonals of the $(n+3)$-gon.
Let~$X(n) \eqdef \set{(a,b)}{0 \le a < b \le n+2}$ (resp.~$Y(n) \eqdef \set{(a,b)}{0 \le a < b-2 \le n}$) denote all (resp.~the internal) diagonals.
The \defn{$\b{g}$-vector} of a diagonal $(a,b) \in X(n)$ is ${\gvector{a,b} \eqdef {\displaystyle\sum}_{\,a < \ell < b} \b{e}_\ell - \frac{b-a-1}{n+1} {\displaystyle\sum}_{\,1 \le \ell \le n+1} \b{e}_\ell}$.
%Note that~$\gvector{a,b} = 0$ if~$(a,b)$ is not internal.
We set~$\gvector{D} \eqdef \set{\gvector{a,b}}{(a,b) \in D}$ for a dissection~$D$.
Recall that:
\begin{compactitem}
\item the set of cones
\(
\gvectorFan[n] \eqdef \bigset{\R_{\ge0} \, \gvectors{D}}{D \text{ dissection of the $(n+3)$-gon}}
\)
forms a complete simplicial fan.
See \cref{fig:lodayFans}\,(left and middle).

@@

\begin{figure}[t]
	\centerline{
	\raisebox{.5cm}{\begin{overpic}[scale=.5]{lodayFan2} \put(0,105){$\boxed{n=2}$} \end{overpic}}
	\hspace{.5cm}
	\begin{overpic}[scale=.5]{lodayFan3} \put(-10,93){$\boxed{n=3}$} \end{overpic}
	\hspace{.5cm}
	\raisebox{.35cm}{\begin{overpic}[scale=.5]{clusterFanA}
		\put(75,92){$\boxed{\B_\circ \! = \!\! {\tiny\left[\begin{array}{@{}r@{}r@{}r@{}} 0 & -1 & 1 \\ 1 & 0 & -1 \\ -1 & 1 & 0 \end{array}\right]}}$}
		\put(35,33){${\scriptstyle x_1}$}
		\put(47,54){${\scriptstyle x_2}$}
		\put(59,33){${\scriptstyle x_3}$}
		\put(70,37){$\frac{x_2 + x_3}{x_1}$}
		\put(22,18){$\frac{x_1 + x_3}{x_2}$}
		\put(29,62){$\frac{x_1 + x_2}{x_3}$}
		\put(80,57){$\frac{x_1 + x_2 + x_3}{x_1 x_3}$}
		\put(35,-7){$\frac{x_1 + x_2 + x_3}{x_1 x_2}$}
		\put(-10,57){$\frac{x_1 + x_2 + x_3}{x_2 x_3}$}
%		\put(34,49){$x_1$}
%		\put(57,61){$x_2$}
%		\put(57,36){$x_3$}
%		\put(68,33){$\frac{x_2 + x_3}{x_1}$}
%		\put(17,36){$\frac{x_1 + x_3}{x_2}$}
%		\put(44,74){$\frac{x_1 + x_2}{x_3}$}
%		\put(90,49){$\frac{x_1 + x_2 + x_3}{x_1 x_3}$}
%		\put(7,12){$\frac{x_1 + x_2 + x_3}{x_1 x_2}$}
%		\put(7,84){$\frac{x_1 + x_2 + x_3}{x_2 x_3}$}
	\end{overpic}}
	}
	\caption{The fans~$\gvectorFan[2]$, $\gvectorFan[3]$ and~$\gvectorFan[\B_\circ]$. The two right $3$-dimensional fans are intersected with the sphere and stereographically projected from the direction~$(-1,-1,-1)$.}
	\label{fig:lodayFans}
	\vspace*{-.2cm}
\end{figure}

@@

For~$\{x,x'\} \in \meshes$ and~$y \in \variables$, we denote by~$\coefficient[y][x][x']$ the coefficient of~$\gvectorFull{\B_\circ}{y}$ in the linear dependence of \cref{lem:dependenceMesh}. % among the $\b{g}$-vectors~$\gvectorFull{\B_\circ}{x}$ and~$\gvectorFull{\B_\circ}{x'}$.
%In other words, by \cref{lem:dependenceMesh}, if~${(\B, \cluster)}$ and~${(\B', \cluster')}$ are two adjacent seeds with~${\cluster \ssm \{x\} = \cluster' \ssm \{x'\}}$, we have~$\coefficient[y][x][x'] = |b_{xy}|$ for~$y \in \cluster \cap \cluster'$ and~$\coefficient[y][x][x'] = 0$ otherwise.
The next statement is proved in~\cite[Sect.~3]{PadrolPaluPilaudPlamondon} using techniques from quiver representation theory.

@@

%We derive from \cref{thm:meshMutations} the following description of the facets of the type cone of the cluster fan.
Using \cref{thm:meshMutations} as a blackbox, we describe the facets of the type cone. See \cref{fig:labelFacetDefiningInequalititiesCA}. %, and thus all polytopal realizations, of the cluster fans of finite type cluster algebras.

\begin{theorem}
\label{thm:extremalExchangeablePairsCA}
For any finite type exchange matrix~$\B_\circ$, the type cone~$\typeCone \big( \gvectorFan[\B_\circ] \big)$ is simplicial and the non-initial mesh mutations are the extremal exchangeable pairs of the cluster fan~$\gvectorFan[\B_\circ]$.
\end{theorem}

@@

A \defn{substring} of a walk~$\smash{\omega = \alpha_1^{\varepsilon_1} \cdots \alpha_\ell^{\varepsilon_\ell}}$ of~$\quiver$ is a string~$\smash{\sigma = \alpha_{i+1}^{\varepsilon_{i+1}} \cdots \alpha_{j-1}^{\varepsilon_{j-1}}}$ of~$\quiver$ for some indices~$1 \le i < j \le \ell$.
The substring~$\smash{\sigma = \alpha_{i+1}^{\varepsilon_{i+1}} \cdots \alpha_{j-1}^{\varepsilon_{j-1}}}$ is \defn{at the bottom} (resp.~\defn{on top}) of the walk~$\omega = \alpha_1^{\varepsilon_1} \cdots \alpha_\ell^{\varepsilon_\ell}$ if~$\varepsilon_i = 1$ and~$\varepsilon_j = -1$ (resp.~if~$\varepsilon_i = -1$ and~$\varepsilon_j = 1$).
In other words the two arrows of~$\omega$ incident to the endpoints of~$\sigma$ point towards~$\sigma$ (resp.~outwards from~$\sigma$).
We denote by~$\Sigma_\bottom(\omega)$ and~$\Sigma_\top(\omega)$ the sets of bottom and top substrings of~$\omega$.

@@

	\caption{The facet-defining inequalities of the type cone~$\typeCone \big( \gvectorFan[\quiver] \big)$ for the non-kissing fan of \cref{fig:nonkissingFans}\,(right). The conventions are similar to \cref{fig:labelFacetDefiningInequalititiesAsso}.}

@@

\begin{definition}
\label{def:brick2acyclic}
A gentle quiver~$\quiver$ is called:
\begin{compactitem}
\item \defn{brick} if any (non necessarily oriented) cycle of~$\quiver$ contains at least two relations in~$I$, %, or equivalently, no walk on~$\quiver$ is self-kissing.
\item \defn{$2$-acyclic} if it contains no cycle of length~$2$.
\end{compactitem}
\end{definition}

@@

We finally consider the graph associahedra studied in~\cite{CarrDevadoss, FeichtnerSturmfels, Postnikov, Zelevinsky}.
Again, we briefly recall all definitions needed here as they are purely combinatorial.

Let~$\graphG$ be a graph with vertex set~$\ground$.
%Let~$\connectedComponents(\graphG)$ be the set of connected components of~$\graphG$ and let $n \eqdef |\ground|-|\connectedComponents(\graphG)|$.
A \defn{tube} of~$\graphG$ is a connected induced subgraph of~$\graphG$.
Let~$\tubes$ denote the set of tubes of~$\graphG$.
The inclusion maximal tubes of~$\graphG$ are its connected components~$\connectedComponents(\graphG)$.
The tubes which are neither empty nor maximal are called \defn{proper}.
Two tubes~$\tube, \tube'$ of~$\graphG$ are \defn{compatible} if they are either nested (\ie $\tube \subseteq \tube'$ or~$\tube' \subseteq \tube$), or disjoint and non-adjacent (\ie $\tube \cup \tube'$ is not a tube of~$\graphG$).
A \defn{tubing} on~$\graphG$ is a set~$\tubing$ of pairwise compatible proper tubes of~$\graphG$.
%See \cref{fig:exmNested}.
The \defn{nested complex} of~$\graphG$ is the simplicial complex~$\nestedComplex(\graphG)$ of all tubings~on~$\graphG$.
